---
output: github_document
format: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(jsonify)
```

# arcgisgeocode

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/arcgisgeocode)](https://CRAN.R-project.org/package=arcgisgeocode)
[![R-CMD-check](https://github.com/R-ArcGIS/arcgisgeocode/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/R-ArcGIS/arcgisgeocode/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

arcgisgeocode provides access to ArcGIS geocoding services from R. It supports address candidate identification, batch geocoding, reverse geocoding, and autocomplete suggestions. 

## Installation

Install the package from CRAN

``` r
# install from CRAN 
install.packages("arcgisgeocode")
```

You can also install the development version from r-universe as a binary for Mac, Windows, or Ubuntu from r-universe like so: 

``` r
# install from R-universe
install.packages("arcgisgeocode", repos = "https://r-arcgis.r-universe.dev")
```

Or you can install the package from source which requires Rust to be available. Follow the  [rustup instructions](https://rustup.rs/) to install Rust and verify your installation is compatible using [`rextendr::rust_sitrep()`](https://extendr.github.io/rextendr/dev/#sitrep). Then install the development version from GitHub:

``` r
# install pak if not available
if (!requireNamespace("pak")) install.packages("pak")

# install development version of {arcgisgeocode}
pak::pak("r-arcgis/arcgisgeocode")
```

## Usage

By default, the [ArcGIS World Geocoder](https://www.esri.com/en-us/arcgis/products/arcgis-world-geocoder) will be used. This geocoding server provides public access to the [`/findAddressCandidates`](https://developers.arcgis.com/rest/geocode/api-reference/geocoding-find-address-candidates.htm), [`/reverseGeocode`](https://developers.arcgis.com/rest/geocode/api-reference/geocoding-reverse-geocode.htm), and [`/suggest`](https://developers.arcgis.com/rest/geocode/api-reference/geocoding-suggest.htm) endpoints made available via the `find_address_candidates()`, `reverse_geocode()`, and `suggest_places()` functions respectively. 

The batch geocoding endpoint [`/geocodeAddresses`](https://developers.arcgis.com/rest/geocode/api-reference/geocoding-geocode-addresses.htm) is available via `geocode_addresses()`. However, this requires the use of an authorization token and may consume credits. 

Refer to the ArcGIS World Geocoder [official documentation](https://developers.arcgis.com/rest/geocode/api-reference/overview-world-geocoding-service.htm) for additional information on use restrictions and licensing. For example, a valid token is required to [store the results](#important-storing-results) of geocoding transactions.

### Reverse geocoding

Reverse geocoding takes a location and finds the associated address. 

::: callout-tip
A token _is not required_ to use this function.
:::

```{r, include = FALSE}
library(pillar)
options("arcgisgeocode.storage" = "never")
```

```{r}
library(arcgisgeocode)

# Find addresses from locations
rev_res <- reverse_geocode(c(-117.172, 34.052))

# preview results
dplyr::glimpse(rev_res)
```

### Address search

The `find_address_candidates()` function returns geocoding candidate results. The function is vectorized over the input and will perform multiple requests in parallel. Each request geocodes one location at a time. 

One or more candidates are returned from the endpoint. You can limit the number of candidates using the `max_locations` argument (with a maximum of 50).

::: callout-tip
A token _is not required_ to use this function.
:::

```{r}
# Find addresses from address search
candidates <- find_address_candidates(
  address = "esri",
  address2 = "380 new york street",
  city = "redlands",
  country_code = "usa",
  max_locations = 2
)

dplyr::glimpse(candidates[, 1:10])
```

### Suggest locations 

Geocoding services can also provide a location suggestion based on a search term and, optionally, a location or extent. The `suggest_places()` function (`/suggest` endpoint) is intended to be used as part of a client-facing application that provides autocomplete suggestions. 

In this example we create a search extent around a single point and find suggestions based on the search term `"bellwood"`. 

::: callout-tip
A token _is not required_ to use this function.
:::

```{r}
# identify a search point as a simple feature column
location <- sf::st_sfc(
  sf::st_point(c(-84.34, 33.74)),
  crs = 4326
)

# buffer and create a bbox object to search within the extent
search_extent <- sf::st_bbox(
  sf::st_buffer(location, 10)
)

# find suggestions within the bounding box
suggestions <- suggest_places(
  "bellwood",
  location,
  search_extent = search_extent
)

suggestions
```

The result is intended to be provided to `find_address_candidates()` to complete the geocoding process. The column `text` contains the address to geocode. The column `magic_key` is a special identifier that makes it much faster to fetch results. Pass this into the argument `magic_key`.

```{r}
# get address candidate information
# using the text and the magic key
res <- find_address_candidates(
  suggestions$text,
  magic_key = suggestions$magic_key
)

dplyr::glimpse(res[, 1:10])
```

## *Important*: Storing results

By default, the argument `for_storage = FALSE` meaning that the results of the geocoding operation cannot be persisted. If you intend to persist the results of the geocoding operation, you must set `for_storage = TRUE`. 

To learn more about free and paid geocoding operations refer to the [storage parameter documentation](https://developers.arcgis.com/documentation/mapping-apis-and-services/geocoding/services/geocoding-service/#storage-parameter). 

## Batch geocoding

Many addresses can be geocoded very quickly using the `geocode_addresses()` function which calls the `/geocodeAddresses` endpoint. Note that this function requires an authorization token. `geocode_addresses()` sends the input addresses in chunks as parallel requests. 

Batch geocoding requires a signed in user. Load the [`{arcgisutils}`](https://github.com/r-arcgis/arcgisutils) to authorize and set your token. This example uses the [Geocoding Test Dataset](# https://datacatalog.urban.org/node/6158/revisions/14192/view) from the [Urban Institute](https://www.urban.org/).

::: callout-tip
A token _is required_ to use this function with the World Geocoding Service. It may not be necessary if you are using a private ArcGIS Enterprise service. 
:::

```{r, eval = FALSE}
library(arcgisutils)
library(arcgisgeocode)

set_arc_token(auth_user())

# Example dataset from the Urban Institute
fp <- "https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/2020/02/25/geocoding_test_data.csv"

to_geocode <- readr::read_csv(fp, readr::locale(encoding = "latin1"))

geocoded <- to_geocode |>
  dplyr::reframe(
    geocode_addresses(
      address = address,
      city = city,
      region = state,
      postal = zip
    )
  )

dplyr::glimpse(res[, 1:10])
```

## Using other locators 

`{arcgisgeocode}` can be used with other geocoding services, including custom locators hosted on ArcGIS Online or Enterprise. For example, we can use the [AddressNC](https://www.nconemap.gov/pages/addresses) geocoding service [available on ArcGIS Online](https://www.arcgis.com/home/item.html?id=247dfe30ec42476a96926ad9e35f725f).

Create a new `GeocodeServer` object using `geocode_server()`. This geocoder can be passed into the `geocoder` argument to any of the geocoding functions. 

```{r}
address_nc <- geocode_server(
  "https://services.nconemap.gov/secure/rest/services/AddressNC/AddressNC_geocoder/GeocodeServer",
  token = NULL
)

res <- find_address_candidates(
  address = "rowan coffee",
  city = "asheville",
  geocoder = address_nc
)

dplyr::glimpse(res[, 1:10])
```
