---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# arcgisgeocode

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/arcgisgeocode)](https://CRAN.R-project.org/package=arcgisgeocode)
[![R-CMD-check](https://github.com/R-ArcGIS/arcgisgeocode/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/R-ArcGIS/arcgisgeocode/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of arcgisgeocode is to provide access to ArcGIS geocoding services from R. Enables address canddiate identification, batch geocoding, reverse geocoding, autocomplete suggestions. 

## Installation

arcgisgeocode uses [`extendr`](https://extendr.github.io/) and requires Rust to be available to install the development version of arcgisgeocode. Follow the  [rustup instructions](https://rustup.rs/) to install Rust and verify your installation is compatible using [`rextendr::rust_sitrep()`](https://extendr.github.io/rextendr/dev/#sitrep).

You can install the development version of arcgisgeocode like so:

``` r
# install pak if not available
if (!requireNamespace("pak")) install.packages("pak")

# install development version of {arcgisgeocode}
pak::pak("r-arcgis/arcgisgeocode")
```

## Example

List available geocoders based on an authorization token. 

```{r}
library(arcgisgeocode)

# create a point
x <- sf::st_sfc(sf::st_point(c(-117.172, 34.052)), crs = 4326)

# reverse geocode
reverse_geocode(x)

# find address candidates
candidates <- find_address_candidates(
  address = c("esri"),
  city = "redlands",
  country_code = "usa"
)

dplyr::glimpse(candidates[,1:10])

# list available geocoding services for signed in account
# list_geocoders()
```
